\documentclass{article}
\begin{document}
encoding - 

\section*{sequences}
quite from thesis - " A network-level coordination of spikes may occur due to a reverberation of the APs or a consistent external drive. The reverberation hypothesis was first proposed by Hebb (1949). According to the this hypothesis APs propagate in a circuit of interconnected cells as in a closed-loop. In a sparse-connectivity network, many closed-loops can be formed on a random basis. Recently, sparsely connected networks became a subject of intensive study (e.g. Amit, 1989). Attractor dynamics are likely to occur within sparsely connected network under certain conditions (Amit, 1989). The network dynamics will converge to a sequence of steady states i.e. to the limit cycle attractor which best supports the coincident input activity to specific cells. More precisely, only those feed-back loops will be stabilized which converge back to the already active cells. According to computer simulations, the conditions necessary for repetition to occur are as follows: (1) spontaneous activity of some cells, (2) fixed refractory periods and (3) feed-back loops in the network (Nadasdy, 1998). The anatomy of the hippocampus CA3 recurrent collateral system is consistent with a sparsely connected network with feed- back loops. In


\section*{place cells - refs}
Gothard 1996 - open field recordings \\
Muller 1994 - the firing rat of place cells are independent of head direction in the open field\\
similar to above : Markus 1995 - the directionality of the place cell depends on the trajectories traversed\\
Wilson MccNaughton 1993 - place cells recorded with tetrode shower fewer subfields\\
Save 1998 - place fields can be also associated with non-visual cues\\
Muller-Kubie 1987 and O'Keefe, Speakman 1987 - Place fields continued to persist after the visual land marks were removed \\
Pivo 1985 - removing some cues did not disrupt the place fields but removing all cues did\\
Quirk, Muller and Kubie 1990 - place fields persisted in the dark when the animals are placed with light and the light turned off.\\
O'keefe, Recce 1993 and Skaags 1996 - spike fired precesses with theta phase as the animal traverses theplace field, also occurs in 2D\\
Mehta, Barenes and Mc Naughton1997 - pfs shift backwards after repeated traversals along the same trajectory 
Place cell firing is also correlated with speed, direction, and turning angle  and stage of the task \\
Markus 1995 - pfs are task sensitive, pfs rapidly changed when the task changed to random foraging and search for food at the corners of a diamond\\
Skaags McNaughton 1996 - found significant correlated temporal bias during SWRs in the SWS after exploration on a linear track.\\
extra hippocampal cells have also been reported, but show different charecteristics the hippocampal place cells\\





\section*{remapping}
O'keefee and burgess 2005- place field firing could be superposition of grid cells, and remappig could be changes ib the offset of grid cells

%%%%%%%%%%%%%%%%%%%%%%%%%%%
It is a capital mistake to theorize before you have all the evidence. It biases the judgment.

Before turning to those moral and mental aspects of the matter which present the greatest difficulties, let the inquirer begin by mastering more elementary problems

You know my method. It is founded upon the observation of trifles.

It is, of course, a trifle, but there is nothing so important as trifles.

...the curious incident of the dog in the night-time... (#9)
     Colonel Ross still wore an expression which showed the poor opinion which he had formed of my companion's ability, but I saw by the inspector's face that his attention had been keenly aroused.
     "You consider that to be important?" he [Inspector Gregory] asked.
     "Exceedingly so."
     "Is there any point to which you would wish to draw my attention?"
     "To the curious incident of the dog in the night-time."
     "The dog did nothing in the night-time."
     "That was the curious incident," remarked Sherlock Holmes.
     The Memoirs of Sherlock Holmes (1893)
     Inspector Gregory and Sherlock Holmes in "Silver Blaze" (Doubleday p. 346-7)
     
     
     

\bibliographystyle{unsrt}
\bibliography{library.bib}
\end{document}