\section{Objectives}
The purpose of this study was chiefly to perform exploratory data analysis leading to the characterization of the spatio-temporal dynamics of the CA3 and CA1 hippocampal networks. This study was aimed at answering some of the  open questions related to how the principal cell population in the hippocampus alter their firing properties and consequently their place field features. How the change of population activity biases the activity during sleep and visa versa, which is an important question directly related to how the internal dynamics and the stimulus related dynamics interact. 
\begin{itemize}
\item The place cell activity changes across multiple environments and the corresponding place field characteristics. 
Whenever a cell is participating in the spatial representation of two or more environments, how is its activity across environments related.          
How would the 
\item Changes in firing properties, if any; that occur in the same environment across multiple sessions.
In a given and unchanging environment, it is interesting to ask weather the firing characteristics of single cells are reliable across multiple sessions and even days.
\item Features of the place cell activity that are preserved in different environments.
Do the place cells that are active in several distinct environments
\item Pairwise firing characteristics of place cells in relation to spatial outlay of their receptive fields.
Once stable place fields have been established, the sequential activation of cells is observed for the place cells that have overlapping spatial receptive fields. When pairs of cells with overlapping place fields are considered, these pairs would show high short time correlation in their firing.
\item The dynamics of population of place cells that are active within and across different environments.
\item Cell sequences in 1 and 2 dimensional arenas that occur during sleep and immobility and their relation to sequences generated during cell sequences occurring when the animal is exploring the environment.
\end{itemize}
 
\st{

Since 2 dimensional environments are closer to the spatial layout in the natural world, the place cells in 2D arena was analyzed in this study. The population activity in the CA3 and CA1 were analysed 


 The place cells in 2D arena was analyzed in this study.}\st{find evidence for the existence of attractors in the hippocampal network. 
search for the computational algorithms used in the hippocampus and its supporting structures for path integration and episodic memory. how well does the data agree with the existing models of path integration and remapping}
