\chapter{Introduction} 

\label{intro}

One \todo{test comment} of the striking features of the brain is its ability to generate self organized activity without the necessary interaction with its of external environment. Although this fact by itself does not imply any functional role played by the self generated activity, it certainly rules out the possibility of completely understanding the brain and its subsystems in a simple stimulus - response paradigm. The task of deducing the mechanisms and computations performed in the sub-structures and  the brain as a whole, would greatly benefit from an approach that examines the brain from an inside-out perspective. This approach would attempt to investigate the internal states that extensively influence the behaviour of the system under scrutiny. The internal states constrain the behaviour of the system when it is interacting with its surroundings and also when the system is isolated from its environment. The internal states might actually be essential for performing the computations required for the appropriate functioning of the system. \\

The oscillations observed in large scale recordings of the brain are one of the chief examples of internal  activity. These oscillatory states are thought to organize the computations at the appropriate time scale. An analogy can be draw between these oscillations and the clock in a microprocessor which organizes the fetching and execution of a sequence of instructions. Although this analogy cannot be taken too far, since the brain is not a sequential processing unit. A set of clocks which tick at different rates seems beneficial for organizing a massive distributed processing system such as the brain. \\

The \emph{Hippocampus} is a structure of the brain that is the focal point of all converging sensory inputs. It presents an excellent window into the internal network dynamics of the brain, far away from the primary somatosensory areas. The hippocampus has a simple and well organised anatomical composition, which provides the opportunity to conduct large scale recordings with fine time resolution. Internal activity is not restricted only to the cortex, the electro-physiology of the hippocampus has revealed several internally self generated patterns. Although during sleep the sensory input is at a minimum, these recordings show a lot of interesting patterns. Some of which are thought to be essential for memory consolidation and spatial processing.  
The hippocampus has come under intensive behaviour combined electro-physiological studies in the context of spatial processing. All these experimental studies address the overarching questions of how the hippocampal network forms spatial representations. To what extent this representation is influenced by the internal dynamics and how this internal activity modulates or is modulated by the incoming sensory stimuli. The examination of these experimental works has revealed a lot about the mechanisms of not just spatial processing, but also has lead to theories postulating general computational architecture of the brain. This study attempts to shed light on general questions related to the degree to which internal states constrain and steer the activity at the network level when it is interacting with its environment and during states with little or no external input. \\
The following sections are intended to be a brief review of the relevant literature and concepts to set the context for this study.
\input{./Chapters/Intro/Sections/anatomy}
\input{./Chapters/Intro/Sections/space}
\input{./Chapters/Intro/Sections/dynamics}
\input{./Chapters/Intro/Sections/objectives}

%\input{./Sections/anatomy}
%\input{./Sections/space}
%\input{./Sections/dynamics}
%\input{./Sections/objectives}

